---
title: "Psychology Students' Reflective Accounts Analysis"
format: 
  revealjs:
    theme: simple
    scrollable: true
    slide-number: true
    code-fold: true
    self-contained: true
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

# Load required libraries
library(tidyverse)
library(quarto)
library(knitr)
library(kableExtra)
library(plotly)
library(wordcloud)
library(tm)
library(sentimentr)

# Read the data
summary_data <- read.csv("data/ReflectiveAccounts_R_Summary.csv")
verbatims <- read.csv("data/ReflectiveAccounts_R_Verbatims.csv")
```

## Introduction

-   Analysis of psychology students' reflections on mini-dissertation experiences
-   Insights into learning journey, challenges, and personal growth
-   Data-driven approach to understanding student experiences

## Data Overview

```{r}
#| label: data-overview
#| echo: false

# Calculate basic statistics
total_accounts <- nrow(summary_data)
avg_sentiment <- mean(summary_data$avg_sentiment, na.rm = TRUE)
unique_themes <- length(unique(summary_data$theme))

# Create a summary table
summary_stats <- data.frame(
  Metric = c("Total Accounts", "Average Sentiment", "Unique Themes"),
  Value = c(total_accounts, round(avg_sentiment, 3), unique_themes)
)

# Display the summary statistics
kable(summary_stats, 
      caption = "Basic Statistics",
      format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

## Sentiment Analysis

```{r}
#| label: sentiment-analysis
#| echo: false

# Create sentiment distribution plot
ggplot(summary_data, aes(x = avg_sentiment)) +
  geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
  theme_minimal() +
  labs(
    title = "Distribution of Sentiment Scores",
    x = "Average Sentiment Score",
    y = "Count"
  )
```

## Theme Analysis

```{r}
#| label: theme-analysis
#| echo: false

# Extract and count themes
theme_counts <- summary_data %>%
  separate_rows(theme, sep = ", ") %>%
  count(theme, sort = TRUE) %>%
  head(10)

# Create theme bar plot
ggplot(theme_counts, aes(x = reorder(theme, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Top 10 Most Common Themes",
    x = "Theme",
    y = "Count"
  )
```

## Word Cloud of Common Terms

```{r}
#| label: wordcloud
#| echo: false

# Create corpus from sample comments
text_corpus <- Corpus(VectorSource(summary_data$sample_comments))

# Clean the text
text_corpus <- tm_map(text_corpus, content_transformer(tolower))
text_corpus <- tm_map(text_corpus, removePunctuation)
text_corpus <- tm_map(text_corpus, removeWords, stopwords("english"))
text_corpus <- tm_map(text_corpus, stripWhitespace)

# Create word cloud
wordcloud(
  text_corpus,
  max.words = 100,
  random.order = FALSE,
  colors = brewer.pal(8, "Dark2")
)
```

## Key Findings

1.  **Sentiment Distribution**
    -   Generally positive sentiment
    -   Most accounts slightly to very positive
    -   Few negative accounts
2.  **Common Themes**
    -   Research design and methodology
    -   Personal growth and learning
    -   Collaboration and project management
3.  **Learning Outcomes**
    -   Growth in research skills
    -   Value of practical experience
    -   Improved time management

## Conclusion

-   Mini-dissertation experience is challenging yet rewarding
-   Students demonstrate significant learning and growth
-   Positive sentiment indicates valuable experience
-   Strong focus on research skills and personal development

## Thank You!

Questions?
